---
title: "Main"
author: "Group work Busschers, Dherbomez, Mather, Van der Meer"
date: "30 Jan 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

<!-- this is the main document generating the report
click the "Knit" button to render the document -->

```{r, echo=FALSE}
#source libraries 
source("libraries.R")

```


#**Part 1.** Introduction to the business project
##Objective 
The objective is to predict hospital capacity need and composition for (area’s in) the Netherlands for the next 25 years.
Broadly, a hospital consists of 6 different parts, which have very different usage, building requirements and costs:

1.	Outpatient beds
2.	Inpatient beds
3.	Outpatient operating rooms
4.	Inpatient operating rooms
5.	Consultation rooms
6.	Imaging diagnostics rooms

##Business use
A number of stakeholders would use these predictions for their decision-making process. Building a new hospital or rebuilding an outmoded hospital is a very expensive process (~200-500M EUR) with a long lead time (from design to finish ~5 years). This brings along significant uncertainty of the future demand. Hospitals are developed for a usage period of about 20-30 years. Therefore, different parties, directly and indirectly, involved in the building and financing of a hospital benefit from better estimations of the future demand for the hospital:

1.	Hospitals: deciding on (re)building a hospital and the size and composition
2.	Banks: assessing investment proposals
3.	Insurers: assessing financing decision for (part of) the build, pricing of provided care, contract negotiations with hospitals
4.	(Local) governments: assessing financing decision for (part of) the build, assessing whether care is sufficiently ensured for the inhabitants

#**Part 2.** Process
We follow a number of predefined steps in this project:

1.	Collect data from the different sources
2.	Clean the data (for example remove double headers)
3.	Do sanity checks and cross check the data against each other (for example, there should be ~17M people in 2017 in the country)
4.	Visualize and summarize the data
5.	Check how data can be matched and which assumptions are needed (for example for outpatient visits, we don’t have information on age and gender, so we will assume that distribution is equal to that of outpatient admissions)
6.	Perform calculations:
    a.	Count number of activities in 6 categories (outpatient admissions, inpatient nursing days, outpatient surgeries, inpatient surgeries, outpatient visits, imaging diagnostics) per region, age and gender
    b.	Multiply number of activities with the corresponding space requirement in m2 to get required m2 per type of space, region in Netherlands, gender and age
    c.	Multiply with demography developments to get required m2 per type of space, region in Netherlands, gender and age for the years 2018-2040.
7.	Visualize results per year and type of space

#**Part 3.** Data
##Description 
All data is publicly available through www.cbs.nl and www.opendis.nl. We will use three different types of data:

1.	Registered care activities per age and gender for the years 2012-2015. This data is in multiple different files: 
    a.	hospital admissions (inpatient and outpatient) per age and gender
    b.	surgeries (inpatient and outpatient) per age and gender
    c.	all activities categorized in 11 categories, not per age and gender. Here fore we will make the assumption that the distribution is equal to that of admissions
2.	Number of square meters needed per building element and single care activity in a year
3.	Forecast of demography of the population (age and gender) for the years 2015-2040.

##Gathering and cleaning
We read our data files. The relevant information has to be extracted and put in a common format

The first dataset is the population forecast per year from 2014 to 2040, per region, type of region(city or land) and gender.
This is how the first lines of the data looks:

```{r, read_data, echo=FALSE}

# population forecast
Pop <- read.csv("Data/CBS_population forecast per region, gender and age - 2014-2040.csv")
visu(Pop,5)
```

The second dataset provides the surface requirement in m² for the various activities that need to be performed 
```{r, echo=FALSE}
# m2 required per activity
BuildReq <- read.csv("Data/Bouwcollege - m2 per activity.csv", header = TRUE, sep=",")
visu(BuildReq,5)
```

the next dataset provides the number of surgeries performed per gender and age between 1995 and 2010. These data required a first cleaning, as it was incomplete. the missing information is replaced by 0 to obtain a clean table that we can work with

```{r, echo=FALSE}
# number of surgeries performed (inpatient and outpatient) per gender and age in 1995-2010
Surgeries <- read.csv("Data/CBS - surgery outpatient and inpatient per gender and age - 1995 - 2010_v2.csv",
header=TRUE, sep=",",dec=".",stringsAsFactors = FALSE, na=0)
 
#cols=c("Total_surgeries","Total_surgeries_per_10000_inhabitants","Inpatient_surgeries","Inpatient_surgeries_per_10000_inhabitants","Outpatient_surgeries","Outpatient_surgeries_per_10000_inhabitants")
 
 
# i=0
# name=0
# for (i in 1:length(cols))
# {
#   name=cols[i]
#   Surgeries$name[Surgeries$name=="."]<-0
# #Surgeries$cols[i][Surgeries$col[i]=="."]<-0
# }
 

visu(Surgeries,5)

Surgeries$Total_surgeries<-as.numeric(as.character(Surgeries$Total_surgeries))
Surgeries$Total_surgeries_per_10000_inhabitants<-as.numeric(as.character(Surgeries$Total_surgeries_per_10000_inhabitants))
Surgeries$Inpatient_surgeries<-as.numeric(as.character(Surgeries$Inpatient_surgeries))
Surgeries$Inpatient_surgeries_per_10000_inhabitants<-as.numeric(as.character(Surgeries$Inpatient_surgeries_per_10000_inhabitants))
Surgeries$Outpatient_surgeries<-as.numeric(as.character(Surgeries$Outpatient_surgeries))
Surgeries$Outpatient_surgeries_per_10000_inhabitants<-as.numeric(as.character(Surgeries$Outpatient_surgeries_per_10000_inhabitants))

Surgeries$Total_surgeries[is.na(Surgeries$Total_surgeries)]<-0
Surgeries$Total_surgeries_per_10000_inhabitants[is.na(Surgeries$Total_surgeries_per_10000_inhabitants)]<-0
Surgeries$Inpatient_surgeries[is.na(Surgeries$Inpatient_surgeries)]<-0
Surgeries$Inpatient_surgeries_per_10000_inhabitants[is.na(Surgeries$Inpatient_surgeries_per_10000_inhabitants)]<-0
Surgeries$Outpatient_surgeries[is.na(Surgeries$Outpatient_surgeries)]<-0
Surgeries$Outpatient_surgeries_per_10000_inhabitants[is.na(Surgeries$Outpatient_surgeries_per_10000_inhabitants)]<-0
               
```

The next dataset indexes some information on hospital admissions per gender and age from 1981 and 2012
```{r, echo=FALSE}
# number of hospital admissions (inpatient and outpatient) per gender and age in 1981-2012
Admissions <- read.csv("Data/CBS - hospital admissions per gender and age 1981-2012_v2.csv", header=TRUE, sep=",")



Admissions$Total_admissions_per_10000_inhabitants<-as.numeric(as.character(Admissions$Total_admissions_per_10000_inhabitants))
Admissions$Outpatient_admissions_per_10000_inhabitants<-as.numeric(as.character(Admissions$Outpatient_admissions_per_10000_inhabitants))
Admissions$Inpatient_admissions_per_10000_inhabitants<-as.numeric(as.character(Admissions$Inpatient_admissions_per_10000_inhabitants))
Admissions$Nursingdays_per_10000_inhabitants<-as.numeric(as.character(Admissions$Nursingdays_per_10000_inhabitants))
Admissions$Average_nursing_days_per_inpatient_admission<-as.numeric(as.character(Admissions$Average_nursing_days_per_inpatient_admission))
Admissions$Average_poulation<-as.numeric(as.character(Admissions$Average_poulation))


Admissions$Total_admissions_per_10000_inhabitants[is.na(Admissions$Total_admissions_per_10000_inhabitants)]<-0
Admissions$Outpatient_admissions_per_10000_inhabitants[is.na(Admissions$Outpatient_admissions_per_10000_inhabitants)]<-0
Admissions$Inpatient_admissions_per_10000_inhabitants[is.na(Admissions$Inpatient_admissions_per_10000_inhabitants)]<-0
Admissions$Nursingdays_per_10000_inhabitants[is.na(Admissions$Nursingdays_per_10000_inhabitants)]<-0
Admissions$Average_nursing_days_per_inpatient_admission[is.na(Admissions$Average_nursing_days_per_inpatient_admission)]<-0
Admissions$Average_poulation[is.na(Admissions$Average_poulation)]<-0


visu(Admissions,5)
```

The next dataset gives the number of details activities per specialisation. Several reference tables are also read to interpret the activities, DOT's and specialisations 

```{r, echo=FALSE}
# number of detailed activities per specialisation, DOT
Act2014 <- read.csv("Data/02_DBC_PROFIEL_2014.csv", header=TRUE, sep=",")

visu(Act2014,5)
```

```{r, echo=FALSE}

# reference tables to interpret the activities, DOT's and specialisations
Ref_Act <- read.csv("Data/03_REF_ZAT.csv", header=TRUE, sep=",")
visu(Ref_Act,5)
Ref_diag <- read.csv("Data/04_REF_DGN.csv", header=TRUE, sep=",")
visu(Ref_diag,5)
Ref_prod <- read.csv("Data/05_REF_ZPD.csv", header=TRUE, sep=",")
visu(Ref_prod,5)
Ref_Spec <- read.csv("Data/06_REF_SPC.csv", header=TRUE, sep=",")
visu(Ref_Spec,5)
```

To ease the data cleaning, we are also loading a technical table which identifies the various age groups and gender, as these are not homogeneous across the raw data
```{r, echo=FALSE}

Age_gender_ID <- read.csv("Data/Age_gender_ID.csv", header=TRUE, sep=",")

visu(Age_gender_ID,5)
```


##Sanity checks
In the graphs below we show how
```{r, sanity_check, echo=FALSE}
#we make a table that summarizes the total number of operations and compare them to what they should be roughly
#we'll cheat by using the final table to the sanity check at the end

summary_table <- data.frame(Type_act= character(6), Number_act_source1=numeric(6), Number_act_source2=numeric(6), Check_value = integer(6))
summary_table$Type_act<- c("Consultation","Imaging diagnostics","Inpatient_surg","Outpatient_surg","Nursing_days","Outpatient_admissions")

# number of activities from OpenDis data write to colum of source 1
# where Zorgprofielklasse_cd 1 = consultation
#       Zorgprofielklasse_cd 2 = Outpatient_admissions
#       Zorgprofielklasse_cd 3 = Nursing days
#       Zorgprofielklasse_cd 4 = Imaging diagnostics
#       Zorgprofielklasse_cd 5 = Inpatient + outpatient surgeries (40%, 60%)

summary_table[which(summary_table$Type_act == "Consultation"),"Number_act_source1"] <- 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="1" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))

summary_table[which(summary_table$Type_act == "Outpatient_admissions"),"Number_act_source1"] <- 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="2" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))

summary_table[which(summary_table$Type_act == "Nursing_days"),"Number_act_source1"] <- 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="3" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))

summary_table[which(summary_table$Type_act == "Imaging diagnostics"),"Number_act_source1"] <- 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="4" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))


summary_table[which(summary_table$Type_act == "Inpatient_surg"),"Number_act_source1"] <- 0.5 * 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="5" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))

summary_table[which(summary_table$Type_act == "Outpatient_surg"),"Number_act_source1"] <- 0.5 * 
  sum(subset(Act2014, (ZORGPROFIELKLASSE_CD=="5" & JAAR=="2014"), select="AANTAL_ZAT", drop=TRUE))

#number of surgeries (inpatient and outpatient), outpatient admissions and nursing days from more detailed CBS source

summary_table[which(summary_table$Type_act == "Inpatient_surg"),"Number_act_source2"] <- sum(subset(Surgeries, Year == "2010", select="Inpatient_surgeries"))

summary_table[which(summary_table$Type_act == "Outpatient_surg"),"Number_act_source2"] <- sum(subset(Surgeries, Year == "2010", select="Outpatient_surgeries"))

Admissions$Outpatient_admissions_total <- Admissions$Outpatient_admissions_per_10000_inhabitants*Admissions$Average_poulation/10000

Admissions$Nursingdays_total<-Admissions$Nursingdays_per_10000_inhabitants*Admissions$Average_poulation/10000

summary_table[which(summary_table$Type_act == "Outpatient_admissions"),"Number_act_source2"] <- sum(subset(Admissions, Year == "2012*", select="Outpatient_admissions_total"))

summary_table[which(summary_table$Type_act == "Nursing_days"),"Number_act_source2"] <- sum(subset(Admissions, Year == "2012*", select="Nursingdays_total"))

summary_table[,"Check_value"]<-c(29400000,12000000,1100000,1400000,9200000,1800000)

#calculate the difference between expected and true value
summary_table$deviation_sc1 <-  abs((summary_table$Number_act_source1-summary_table$Check_value)/summary_table$Check_value)
summary_table$deviation_sc2 <- abs((summary_table$Number_act_source2-summary_table$Check_value)/summary_table$Check_value)
summary_table[1:2,"deviation_sc2"]<-0

# make a plat of it:  Simple Horizontal Bar Plot with Added Labels 
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin.
barplot(summary_table$deviation_sc1, main="Deviation of source OpenDis from expected values", horiz=TRUE, names.arg=c("Consultation","Imaging diagnostics","Inpatient surgery","Outpatient surgery","Nursing days","Outpatient admissions"),col="red2")

par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin.
barplot(summary_table$deviation_sc2, main="Deviation of source OpenDis from expected values", horiz=TRUE, names.arg=c("Consultation","Imaging diagnostics","Inpatient surgery","Outpatient surgery","Nursing days","Outpatient admissions"),col="red2")
```

##Summary and visualisation
To show how the demography of the population changes, we visualize the composition of the population in 2014 and 2040.

```{r, visualize_pop, echo=FALSE}


pop_graph <- subset(Pop, (Age=="0 tot 5 jaar" |  Age=="5 tot 10 jaar"|  Age=="10 tot 15 jaar"|  Age=="15 tot 20 jaar"|  Age=="20 tot 25 jaar"|  Age=="25 tot 30 jaar"|  Age=="30 tot 35 jaar"|  Age=="35 tot 40 jaar"|  Age=="40 tot 45 jaar"|  Age=="45 tot 50 jaar"|  Age=="50 tot 55 jaar"|  Age=="55 tot 60 jaar"|  Age=="60 tot 65 jaar"|  Age=="65 tot 70 jaar"|  Age=="70 tot 75 jaar"|  Age=="75 tot 80 jaar"|  Age=="80 tot 85 jaar"|  Age=="85 tot 90 jaar"|  Age=="90 tot 95 jaar"|  Age=="95 jaar of ouder"),select=c(Age, Population_x1000, Type.of.region, Gender, Year), drop = TRUE)


pop_graph_men_2014 <- subset(pop_graph, (Type.of.region=="Country" & Year == "2014" & Gender=="Mannen"),select=c(Population_x1000,Age), drop = TRUE)

pop_graph_women_2014 <- subset(pop_graph, (Type.of.region=="Country" & Year == "2014" & Gender=="Vrouwen"),select=c(Population_x1000, Age), drop = TRUE)

pop_graph_men_2040 <- subset(pop_graph, (Type.of.region=="Country" & Year == "2040" & Gender=="Mannen"),select=c(Population_x1000,Age), drop = TRUE)

pop_graph_women_2040 <- subset(pop_graph, (Type.of.region=="Country" & Year == "2040" & Gender=="Vrouwen"),select=c(Population_x1000,Age), drop = TRUE)

Pop_graph_2014<- merge(pop_graph_men_2014, pop_graph_women_2014, by= "Age")
Pop_graph_2014_ordered <- Pop_graph_2014[,c(2,3,1)]

pyramid(Pop_graph_2014_ordered, Laxis=NULL, Raxis=NULL,
AxisFM="g", AxisBM="", AxisBI=3, Cgap=0.6, Cstep=0.5, Csize=1,
Llab="Males", Rlab="Females", Clab="Ages", GL=TRUE, Cadj=-0.03,
Lcol="Cyan", Rcol="Pink", Ldens=-1, Rdens=-1, main="Population 2014")

Pop_graph_2040<- merge(pop_graph_men_2040, pop_graph_women_2040, by= "Age")
Pop_graph_2040_ordered <- Pop_graph_2040[,c(2,3,1)]

pyramid(Pop_graph_2040_ordered, Laxis=NULL, Raxis=NULL,
AxisFM="g", AxisBM="", AxisBI=3, Cgap=0.6, Cstep=0.5, Csize=1,
Llab="Males", Rlab="Females", Clab="Ages", GL=TRUE, Cadj=-0.03,
Lcol="Cyan", Rcol="Pink", Ldens=-1, Rdens=-1, main="Population 2040")

```

#**Part 4.** Calculations
##Preparation of data
As visualized above, the raw data that we are getting from the various sources are not homogeneous. we need to organize and combine them in order to be able to use them together and draw the projections up to 2040.

we start by adding the Age/gender ID to the population forecast table to be able to perform projections later on

```{r,add_ID_age_gender,echo=FALSE}

# add Index numbers to population table and sum total within ID category
Pop_index_temp <- subset(merge(Pop, Age_gender_ID, by.x=c("Age", "Gender"), by.y=c("Forecast_age", "Gender")),select=c("Age_gender_ID","Year", "Region", "Type.of.region","Population_x1000"))

Pop_index <- aggregate(Population_x1000 ~ .,Pop_index_temp, sum)

```


The surgeries contain information for two of our six defined usages: inpatient and outpatient operating rooms (surgeries). For each gender or age group, we index the number of activities performed per 1000 inhabitant and the required corresponding surface.

```{r, SurgeriesTable, echo=FALSE}
# select all rows for which the column surgeries says 'all', year=2010
# select columns containing 1 age, 2 gender, 3 number of inpatient surgeries and 4 number of outpatient surgeries
Surgeries_2 <- subset(Surgeries, (Surgery=="All"& Year=="2010"), select=c(Gender,Age,Inpatient_surgeries,Outpatient_surgeries),drop=TRUE)

# add age_gender_ID
Surgeries_3 <- unique(subset(merge(Surgeries_2, Age_gender_ID, by.x=c("Age", "Gender"), by.y=c("Surg_Age", "Gender")),select=c("Age_gender_ID","Age","Gender","Inpatient_surgeries","Outpatient_surgeries")))

# restructure table such that it contains a column for type of activity, instead of each column being a different activity
Surgeries_inp <- Surgeries_3[,c("Age_gender_ID","Age", "Gender", "Inpatient_surgeries")]
colnames(Surgeries_inp) <- c("Age_gender_ID","Age", "Gender", "Number_of_act") # rename column
Surgeries_inp$Type_act="Inpatient_surg" # add column

Surgeries_outp <- Surgeries_3[,c("Age_gender_ID","Age", "Gender", "Outpatient_surgeries")]
colnames(Surgeries_outp) <- c("Age_gender_ID","Age", "Gender", "Number_of_act") # rename column
Surgeries_outp$Type_act="Outpatient_surg" # add column

# add total number of inhabitants per Age_gender_ID in two seperate files for inpatient and outpatient surgeries

# merge surgeries with a subset of Popindex that contains the data for 2014 and NL
# note: shortcut of using 2014 population against 2010 numbers, because 2010 is not in the table
Surgeries_def_inp <- merge(Surgeries_inp,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID")

Surgeries_def_inp$Act_per_1000inhab <- as.numeric(Surgeries_def_inp$Number_of_act) / as.numeric(Surgeries_def_inp$Population_x1000)

# add square meter requirement
Surgeries_def_inp$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Surgeries"),"Total.normative.space"]

Surgeries_def_outp <- merge(Surgeries_outp,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID")

Surgeries_def_outp$Act_per_1000inhab <- as.numeric(Surgeries_def_outp$Number_of_act) / as.numeric(Surgeries_def_outp$Population_x1000)

# add square meter requirement
Surgeries_def_outp$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Surgeries"),"Total.normative.space"]


visu(Surgeries_def_inp,5)
visu(Surgeries_def_outp,5)

```

The raw data with the admissions give information on the inpatient beds (the nursing days) and outpatient beds per patient and for each kind of medical act being performed being performed. We work the data in order to obtain a table with one column indexing all the different types of activities and their occurence, per age and gender type. This data. The data is extrated for the year 2012 only, which is the only one we want to work with (because the most recent).

```{r, AdmissionsTable, echo=FALSE}
# select all rows for which the column year=2012
# select columns containing 1 age, 2 gender, 3 number of outpatient admissions per 10000 inhabitants,  4 number of nursing days per 10000
Admissions_2 <- subset(Admissions, Year=="2012*", select=c(Gender,Age,Outpatient_admissions_per_10000_inhabitants, Nursingdays_per_10000_inhabitants, Average_poulation),drop=TRUE)

# add age_gender_ID
Admissions_3 <- unique(subset(merge(Admissions_2, Age_gender_ID, by.x=c("Age", "Gender"), by.y=c("Adm_age", "Gender")),select=c("Age_gender_ID","Surg_Age","Gender","Outpatient_admissions_per_10000_inhabitants", "Nursingdays_per_10000_inhabitants", "Average_poulation")))
colnames(Admissions_3)<-c("Age_gender_ID","Age","Gender","Outpatient_admissions_per_10000_inhabitants", "Nursingdays_per_10000_inhabitants", "Average_poulation")

# restructure table such that it contains a column for type of activity, instead of each column being a different activity
Adm_inp <- Admissions_3[,c("Age_gender_ID","Age", "Gender", "Nursingdays_per_10000_inhabitants", "Average_poulation")]

# weighted average to find nursing days in a category, because we want to use bigger categories later
Adm_inp$Number_of_act = Adm_inp$Nursingdays_per_10000_inhabitants * Adm_inp$Average_poulation/10000
Adm_inp2 <- Adm_inp[,c("Age_gender_ID","Age", "Gender","Number_of_act")]
Adm_inp3 <- aggregate(Number_of_act ~ .,Adm_inp2, sum)                                  
                                  
Adm_inp3$Type_act="Nursing_days" # add column
Adm_inp3$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Nursing days"),"Total.normative.space"]

Adm_outp <- Admissions_3[,c("Age_gender_ID","Age", "Gender", "Outpatient_admissions_per_10000_inhabitants", "Average_poulation")]

Adm_outp$Number_of_act = (as.integer(Adm_outp$Outpatient_admissions_per_10000_inhabitants) * Adm_outp$Average_poulation)/10000
Adm_outp2 <- Adm_outp[,c("Age_gender_ID","Age", "Gender","Number_of_act")]
Adm_outp3 <- aggregate(Number_of_act ~ .,Adm_outp2, sum)  

Adm_outp3$Type_act="Outpatient_admissions" # add column
Adm_outp3$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Day admissions"),"Total.normative.space"]

# add total number of inhabitants per Age_gender_ID

# merge surgeries with a subset of Popindex that contains the data for 2014 and NL
# note: shortcut of using 2014 population against 2010 numbers, because 2010 is not in the table
Admissions_def_inp <- merge(Adm_inp3,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID")

Admissions_def_inp$Act_per_1000inhab <- as.numeric(Admissions_def_inp$Number_of_act) / as.numeric(Admissions_def_inp$Population_x1000)

Admissions_def_inp <- Admissions_def_inp[,c("Age_gender_ID","Age","Gender","Number_of_act","Type_act","Population_x1000","Act_per_1000inhab","Space_per_act")]

Admissions_def_outp <- merge(Adm_outp3,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID")

Admissions_def_outp$Act_per_1000inhab <- as.numeric(Admissions_def_outp$Number_of_act) / as.numeric(Admissions_def_inp$Population_x1000)

Admissions_def_outp <- Admissions_def_outp[,c("Age_gender_ID","Age","Gender","Number_of_act","Type_act","Population_x1000","Act_per_1000inhab","Space_per_act")]


visu(Admissions_def_inp,5)
visu(Admissions_def_outp,5)

```

The last two activities we want insights on are the consultation and imaging diagnostics.The information is included in the raw data about the outpatient surgeries. we only want the data for the most recent year.

```{r echo=FALSE} 
Consultation_imaging_ref <- subset(Surgeries_outp, select=c("Age_gender_ID","Age","Gender","Number_of_act"), drop=TRUE)

Consultation_imaging_ref$Ratio_per_agegroup= as.integer(Surgeries_outp$Number_of_act)/sum(as.integer(Surgeries_outp$Number_of_act))  # add column with ratio per age-group

Consultation_imaging_ref $Type_Act_con="Consultation" # add column for type of activity (consultations)

Consultation_1 <- subset(Act2014, (ZORGPROFIELKLASSE_CD=="1" & JAAR=="2014"), select=c("JAAR","AANTAL_ZAT"), drop=TRUE) # Get total number of activities for this category (consultations)

Consultation_imaging_ref$Activities_con=round( (Consultation_imaging_ref$Ratio_per_agegroup)*sum(Consultation_1[,2]),0) # Create proxy for number of activities per agegroup and gender based on Outpatient ratio 

Consultation_imaging_ref$Type_Act_img="Imaging diagnostics" #  add column for type of activity (Imaging)

Imaging_1 <- subset(Act2014, (ZORGPROFIELKLASSE_CD=="7" & JAAR=="2014"), select=c("JAAR","AANTAL_ZAT"), drop=TRUE) # Get total number of activities for this category (Imaging)

Consultation_imaging_ref$Activities_img=round( (Consultation_imaging_ref$Ratio_per_agegroup)*sum(Imaging_1[,2]),0) # Create proxy for number of activities per agegroup and gender based on Outpatient ratio

Consultation_2 <- subset(Consultation_imaging_ref, select=c("Age_gender_ID","Age","Gender","Activities_con","Type_Act_con"), drop=TRUE) # Prepare data for output table to be used in calculations

colnames(Consultation_2) <- c("Age_gender_ID","Age", "Gender", "Number_of_act","Type_act") # rename column # rename columns in order to match with other type of activities

Imaging_2 <- subset(Consultation_imaging_ref, select=c("Age_gender_ID","Age","Gender","Activities_img","Type_Act_img"), drop=TRUE)

colnames(Imaging_2) <- c("Age_gender_ID","Age", "Gender", "Number_of_act","Type_act") # rename column # rename columns in order to match with other type of activities

Consultation_def <- merge(Consultation_2,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID") # add population numbers

Consultation_def$Act_per_1000inhab <- as.numeric(Consultation_def$Number_of_act) / as.numeric(Consultation_def$Population_x1000) # calculate activities per 1000 inhabitants

Consultation_def$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Outpatient visits"),"Total.normative.space"] # add square meter requirement

Imaging_def <- merge(Imaging_2,subset(Pop_index, (Type.of.region=="Country"& Year=="2014"), select=c(Age_gender_ID,Population_x1000),drop=TRUE), by="Age_gender_ID") # add population numbers

Imaging_def$Act_per_1000inhab <- as.numeric(Imaging_def$Number_of_act) / as.numeric(Imaging_def$Population_x1000) # calculate activities per 1000 inhabitants

Imaging_def$Space_per_act = BuildReq[which(BuildReq$Englisch.name.activity.cluster == "Imaging diagnostics"),"Total.normative.space"] # add square meter requirement

visu(Consultation_def,5)
visu(Imaging_def,5)

```
We finally obtain a table that we can work with, indexing total activites per 1000 inhabitant, per age, gender, region and that we will take as reference for the year 2014, allowing us to then project the future surface needs per activities from 2012 to 2040.
```{r echo=FALSE}

Total_activities_2014 <- rbind.fill(Admissions_def_outp,Surgeries_def_outp,Surgeries_def_inp, Consultation_def, Admissions_def_inp,Imaging_def)
visu(Total_activities_2014,10)


```



##Assumptions
[...write about the assumptions that we have to make..]

##Implementation

To build our projection, we are combining the data from the population forecast and the activities table for 2014. We basically perform the following steps:
a.	Count number of activities in 6 categories (outpatient admissions, inpatient nursing days, outpatient surgeries, inpatient surgeries, outpatient visits, imaging diagnostics) per region, age and gender

b.	Multiply number of activities with the corresponding space requirement in m2 to get required m2 per type of space, region in Netherlands, gender and age

c.	Multiply with demography developments to get required m2 per type of space, region in Netherlands, gender and age for the years 2018-2040.


```{r, Projection, eval=TRUE, echo= FALSE}
cproj=c("Year", "Region", "Activity")
#Generating the global table with projection of space from 2014 to 2040

Projection <- merge(subset(Pop_index, select=c(Age_gender_ID, Year, Region, Population_x1000 ),drop=TRUE), subset(Total_activities_2014, select=c(Age_gender_ID, Age,Gender, Type_act, Act_per_1000inhab, Space_per_act),drop=TRUE), by="Age_gender_ID")
 
Projection$Space_x1000m2=round((Projection$Population_x1000)*(Projection$Act_per_1000inhab)* (Projection$Space_per_act)/1000,0) # Calculate total space needed for every activity in every agegroup for both gender 

#subset per year, age group, gender, region and activity

Projection<- subset(Projection,select = c(Age,Gender,Year,Region,Type_act,Space_x1000m2))

colnames(Projection)[colnames(Projection) == 'Type_act'] <- 'Activity' 
visu(Projection,20)
```

Now that we have the detailed projections, we want sum the required surfaces on the Age and Gender to obtain statistics pe Region and Activity
```{r, result, eval=TRUE, echo=FALSE}


Result <-aggregate(Space_x1000m2~ Year+Region+Activity,Projection,sum)
 
visu(Result,20)
```



#**Part 5.** Results
[...table and visualisation of results...]
[...some textual explanation of it...]

```{r} 


test<- matrix(c(2014,2015,2016,2014,2015,2016,"Diagnostics","Diagnostics","Diagnostics","Surgeries","Surgeries","Surgeries",10,12,19,11,13,17),nrow=6, ncol=3)
colnames(test)<-c("Year", "Group", "Activity")
test_graph <- data.frame(test)
ggplot(test_graph, aes(x=Year,y=Activity,group=Group,fill=Group)) + geom_area() + geom_line(position='stack')

library(ggplot2)
Result_graph <- data.frame(subset(Result, (Region=="Nederland"), select=c("Year","Activity","Space_x1000m2"), drop=TRUE) )
ggplot(Result_graph, aes(x=Year,y=Space_x1000m2,group=Activity,fill=Activity),order(Result_graph$Activity)) + geom_area(stat="identity") + geom_line(position='stack') + labs(title="Needed space per category in m2")
Result_graph2 <- data.frame(subset(Result_graph_ordered, (Year==2015|Year==2020| Year==2025|Year==2030|Year==2035|Year==2040),select=c("Year","Activity","Space_x1000m2"), drop=TRUE))
ggplot(Result_graph2, aes(x = Year, y = Space_x1000m2, fill = Activity, label = Space_x1000m2)) +
  geom_bar(stat = "identity") +
  geom_text(size = 2, position = position_stack(vjust = 0.2))

```


#**Part 6.** Usage of process and results

##Limitations
There are a number of limitations to the described process for aplication in real-life business problems, mainly caused by availibility of data:

1.  We don’t know the current capacity, its usage or its life expectancy. So we cannot match the required space to the actual space to determine how much extra should be build for the future.
2.  We assume no impact of innovation of care. However, in practice we have seen that nursing days per admission has steadily decreased. The model could be refined to include assumptions on this.
3.  We have no data on specific zip-codes. Including makes it possible to forecast the space requirement for a specific hospital based on its care region.

##Usage in different context
We can use the same process for different data sets and purposes:

1.  Different time horizons
2.  Different countries 
3.  Refining for specific parts of the country, for example the care region of a hospital
4.  Refining for specializations
5.  Different types of care, such as youth care. Activities are different and space requirements as well, but the process is the same
6.  Alternative ‘weighting’ of activities. By using m2 to weight the 6 different types of care, we get a physical result, but we could also weight the activities with average cost to forecast future health expenditure.
